Preregistration for modelling in ecolgy and conservation

### 2020-09-12 index.Rmd
revise working draft - restructure user design section

---
title: "Preregistration for modelling in ecolgy and conservation"
description: |
  Working Draft.
author:
  - name: Elliot Gould 
    affiliation: School of BioSciences, University of Melbourne
  - name: Chris Jones
    affiliation: Department of Environment Land and Water
  - name: Libby Rumpff
    affiliation: School of BioSciences, University of Melbourne
  - name: Hannah Fraser
    affiliation: School of BioSciences, University of Melbourne
  - name: Fiona Fidler
    affiliation: School of BioSciences, University of Melbourne
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    self_contained: false
---
**Intended journal:** Methods in Ecology and Evolution.

@@

### User-centred design approach ###

**Designing Preregistration Templates with a User-Centred Design Approach**

We took a user-centred design (UCD) approach to the task of translating preregistration into model-based research contexts for ecology and conservation. UCD is a philosophy for designing services or products that accounts for user requirements at all stages of the design process --- the goal of which is to generate both an effective and usable tool for specific people in specific contexts [@Pavelin2012]. As Pavelin et al illustrate,  it is critically important that users be involved in the design process because it is difficult to predict how users will actually use a resource. This mirrors Pu et al's qualitative analysis into preregistration design, that a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings. Consequently, although user-centered design is typically used within software engineering domains, we heed Pu's advice that "It is crucial to treat preregistration [...] as a user-centered design problem to ensure preregistration achieves its intended goal."

We borrow from the UCD methodology proposed by Pavelin et al. [-@Pavelin2012], and follow the core steps ![](./figures/user-centered-design-preregistration.png): 1. Define the goals of the resource under development and identify audience, 2. Characterise users and their needs, 3. Design mock-ups of your product, 4. Build prototypes of selected mock-ups and test with users, 5. Implement the product, writing technical implementations and building the product. In the case of this project, the 'product' consists of a preregistration template, specification of the internal logic of the template, including an updated domain-specific procedure for using the template, a user-interface for completing the preregistration, and a user-guide to accompany the template and interface.

#### Define Vision & Objectives ####

@@

Defining the purpose and intended use-cases of a preregistration template has important implications for the design of the template --- from its structure, content, user-interface and use within the broader research workflow. We define the primary purpose of the template as "to delimit researcher degrees of freedom" when conducting model-based research in applied ecology and conservation. Although others have defined the purpose of preregistration to include 'delineating between *exploratory* from *confirmatory* research' \<refs\>, this is not the intended purpose of the template here. Beyond hypothesis-testing contexts, this is a simplifying dichotomy that, in practice, is more blurred and nuanced than the ideal. For instance researchers might move back and forth between different reasoning-modes throughout the modelling cycle. Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, and consideration of research from philosophy of science. Consequently, we wish to deemphasise the purpose of distinguishing between exploratory and confirmatory research, and focus the design of the templates in being able to delimit researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and other questionable research practices, like cherry-picking, HARKing etc. \<insert ref to my QRP manuscript\>.

@@

Given the stated purposes of the preregistration templates we designed, it implies two types of users: creators of preregistrations using the template, and reviewers of a completed template, including reviewers, journal editors and any readership. The content and structure of the items on the template will be designed with the preregistration-creator and the preregistration readership in mind - it will capture critical decision-points that may influence the overall model and analysis and recommendations or decisions that the model informs, thus delimiting researcher degrees of freedom at these junctures in the modelling decision process. It should also enhance the research transparency of the study, ensuring that these choices are documented with sufficient transparency to let the reader make informed decisions about the credibility and reliability of the model and associated analysis and interventions.

@@

**Use-Cases**

@@

#### User Research ####

On the 4^th^ March 2020, we ran a series of structured activities designed to inform the development of two pre-registration templates, one for field-work and sampling design, the other for modelling and data-analysis. We split the activities into two groups, to better focus discussion for each of these templates. All workshop materials and workshop outputs can be found on the OSF component here \<LINK\>.

@@

**Activity 1 - Capturing Scientific Workflows and Decision Points**

The objective of this activity was to capture critical decision-steps in the design of both field sampling efforts and modelling exercises. This activity was intended to identify key decision-steps in the scientific process that should be included in a preregistration template.  The first part of this activity was done as individuals, where we reflected on our scientific process for a recent or memorable research project. We listed the steps in the workflow from beginning to end. We attempted to document  critical decision-steps in this process that should be present on a preregistration template. By "critical" we mean that without these decision-steps the template would not a) adequately constrain researcher degrees of freedom and b) transparently document the analytic decision-process.

In the second part of this exercise, we came together as a group to collate each of the steps in our personal workflows onto a group workflow template This template was prepared earlier (\<specific OSF link to the document\>), and consisted of several key 'phases' of the scientific process, with decision-steps grouped under each phase. The development of the draft modelling workflow was derived from previous work by Gould et al. *in prep.,* whilst the field sampling workflow was created specifically for this exercise. We sorted and categorised each decision-step as belonging to a particular workflow phase. Next we grouped similar decision-steps that could be included as a single item on a preregistration template. At the end of this process we reviewed the suggested phases on the workflow to assess whether phases needed further discretisation, aggregation, removal, or amendment.

@@

This second activity aimed to identify critical challenges and roadblocks to the implementation and uptake of preregistration in ecology and conservation, as well as their solutions. We conducted two lots of structured discussions, one for each of the field and modelling template groups. Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature. These talking points were chosen with the aim of identifying problems in the content and or structuring of the phases and decision steps that were collectively identified in activity 1, as well as any conceptual or philosophical issues with preregistration in ecology, conservation, and in particular modelling, that might impact on the template structure, content and procedural implementation of preregistration.

##### Workshop Analysis #####

#### Design ####

##### Translating Workflows into a Preregistration Template #####

**1\. Synthesise Workshop Outputs: Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**

@@

A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on 'good modelling practice' and structured decision-making literature in ecological management. Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substep' \<insert link to appendices or just present the final?\>. Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and substeps. We revised the coding structure \<inductively (is this the word)\>, according to patterns and themes identified across multiple personal workflows. For example, we added the final phase \<name for last phase\>, which aimed to capture the fact that analysts usually present the results of their modelling to clients, decision-makers or other stakeholders that must interpret the evidence and are responsible for making management decisions. In this way, we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both \<idealised norms and norms of practice, Pu et al., see other philosophy of sci refs\>. This workflow is used to provide structure and content to the preregistration template.

@@

We translated the model-based workflow identified in part I of workshop analysis into a text-based preregistration template. Reflecting on existing templates \<link to OSF repository of templates \>, we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part I of workshop output analysis. We synthesised the following ontologies:

@@

2\. Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps'. Corresponds to an activity or set of activities in a scientific workflow. \<again, these are being scrapped and turned into tool-tips.\>

@@

[@Moallemi2019] featuresof a good guideline:

a) A standard step-wise structure relevant to a variety of (exploratory modelling) methods in robust decision

support.

(b) The enumeration of (methodological and mixing design) choices available at each step to reflect on alternative

options available to researchers.

(c) Examples/illustrations to clarify the step of the guideline and choices at each step.

(d) An easy to follow checklist to promote wider adoption.

@@

@@

##### Redesigning a new methodology for preregistering model-based ecology and conservation research #####

##### 'In-situ' Evaluation and Testing of Templates #####

@@

Failing to make the 'internal model of preregistration' clear to users is a critical source of confusion among preregistration template users that may impede preregistration uptake, and undermine the ability of the template to achieve its objectives (Pu et al. 2020). Consequently, in this section we articulate the inner logic of our preregistration templates that informs the the content, structure of the template, and our proposed procedure and interface for implementing the template.

@@

This will be a box or a table with numbered items, with just the preregistration item directives or question. The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'. Could possibly go into the appendix, but I think given that it's the primary research output, it should go into the main-text of the manuscript.

@@

I thinkthe preregistration can easily be usd with a reporting check

>>>>

---
title: "Preregistration for modelling in ecolgy and conservation"
description: |
  Working Draft.
author:
  - name: Elliot Gould 
    affiliation: School of BioSciences, University of Melbourne
  - name: Chris Jones
    affiliation: Department of Environment Land and Water
  - name: Libby Rumpff
    affiliation: School of BioSciences, University of Melbourne
  - name: Hannah Fraser
    affiliation: School of BioSciences, University of Melbourne
  - name: Fiona Fidler
    affiliation: School of BioSciences, University of Melbourne
date: "`r Sys.Date()`"
bibliography: ./data/references.bib
output: 
  distill::distill_article:
    self_contained: false
---
**Intended journal:** Methods in Ecology and Evolution.

@@

### Designing Preregistration Templates with a User-Centred Design Approach ###

We took a user-centred design (UCD) approach to the task of translating preregistration into model-based research contexts for ecology and conservation. UCD is a philosophy for designing services or products that accounts for user requirements at all stages of the design process --- the goal of which is to generate both an effective and usable resource for specific people in specific contexts [@Pavelin2012]. As Pavelin et al illustrate,  it is critically important that users be involved in the design process because it is difficult to predict how users will actually use a resource. This mirrors Pu et al's qualitative analysis of preregistration template design, finding that a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings. Consequently, although user-centered design is typically used within software engineering domains, we heed Pu's advice that "It is crucial to treat preregistration [...] as a user-centered design problem to ensure preregistration achieves its intended goal."

We borrow from the UCD methodology proposed by Pavelin et al. [-@Pavelin2012], and follow the core steps ![](./figures/user-centered-design-preregistration.png):

1\. Define the goals of the resource under development and identify the audience,

2\. Characterise users and their needs,

3\. Design mock-ups of your product,

4\. Build prototypes of selected mock-ups and test with users,

5\. Implement the product, write technical implementations and build the product.

In this case, the 'product' so-to-speak, consists of a preregistration template, specification of the internal logic of the template, including an updated domain-specific procedure for using the template, a user-interface for completing the preregistration, and a user-guide to accompany the template and interface.

The first step in a user-centred design approach is to articulate the goals, objectives and target audience of the tool, consequently we define the intended purpose of the preregistration, the intended users of the templates, and their research contexts. These aspirations for the preregistration template are the guiding principles for the development and evaluation of ecology-specific preregistration, and have important implications for the design of the template - from its structure, content, user-interface and usage within the broader research workflow.

@@

We define the primary purpose of the template as "to delimit researcher degrees of freedom" when conducting model-based research in applied ecology and conservation. Although others have defined the purpose of preregistration to include 'delineating between *exploratory* from *confirmatory* research' \<refs\>, this is not the intended purpose of the template here. Beyond hypothesis-testing contexts, this is an idealised and simplifying dichotomy that, in practice, is more complicated and nuanced. For instance researchers might move back and forth between different reasoning-modes throughout the modelling cycle. Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, and consideration of perspectives from philosophy of science. Consequently, we wish to deemphasise the purpose of distinguishing between exploratory and confirmatory research, and focus the design of the templates in being able to delimit researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and other questionable research practices, like cherry-picking, HARKing etc. \<insert ref to my QRP manuscript\>.

@@

Given the stated purposes of the preregistration templates we designed, it implies two types of users: creators of preregistrations using the template, and reviewers of a completed template, including reviewers, journal editors and any readership. The content and structure of the items on the template will be designed with the preregistration-creator and the preregistration readership in mind - it will capture critical decision-points that may influence the overall model and analysis and recommendations or decisions that the model informs, thus delimiting researcher degrees of freedom at these junctures in the modelling decision process. It should also enhance the research transparency of the study, ensuring that these choices are documented with sufficient transparency to let the reader make informed decisions about the credibility and reliability of the model and associated analysis and interventions. (MOVE RESEARCH TRANSPARENCY QUOTES FROM BELOW TO HERE).

@@

**Preregistration use-cases**

@@

On the 4^th^ March 2020, we ran a series of structured activities designed to inform the development of two preregistration templates, one for field-work and sampling design, the other for modelling and data-analysis. We split the activities into two groups, to better focus discussion for each of these templates. All workshop materials and workshop outputs can be found on the OSF component here \<LINK\>.

@@

Failing to make the 'internal model of preregistration' clear to users is a critical source of confusion among preregistration template users that may impede preregistration uptake, and undermine the ability of the template to achieve its objectives (Pu et al. 2020). Consequently, we aim to articulate a generalised modelling workflow that is used to structure the preregistration template, and to guide the selection of items for inclusion in the template.

## Workshop Activity 1: Identifying Ecological Modelling Workflows Norms and Practice

The objective of this activity was to identify modelling workflows that capture critical decision-steps in the process of field sampling efforts and modelling exercises that should be included in a preregistration template. By "critical" we mean that without these decision-steps the template would not a) adequately constrain researcher degrees of freedom and b) transparently document the analytic decision-process. The activity consisted of two-parts, first, we individually reflected on our scientific process for a recent or memorable research project, listing the steps we undertook from the beginning to end.  In the second part of this exercise, we came together as a group to collate each of the steps in our personal workflows onto a group workflow template. This template was prepared earlier (\<specific OSF link to the document\>), and consisted of several key 'phases' of the scientific process, with decision-steps grouped under each phase. The development of the draft modelling workflow was derived from previous work by Gould et al. *in prep.,* whilst the field sampling workflow was created specifically for this exercise. We sorted and categorised each decision-step as belonging to a particular workflow phase. Next we grouped similar decision-steps that could be included as a single item on a preregistration template. At the end of this process we reviewed the suggested phases on the workflow to assess whether phases needed further discretisation, aggregation, removal, or amendment.

@@

This second activity aimed to identify critical challenges and roadblocks to the implementation and uptake of preregistration in ecology and conservation, as well as brainstorming solutions to these challenges. We conducted two lots of structured discussions, one for each of the field and modelling template groups. Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature \<cite / link to OSF, or even discuss??\>. These talking points were chosen with the aim of identifying problems in the content and or structuring of the phases and decision steps that were collectively identified in activity 1, as well as any conceptual or philosophical issues with preregistration in ecology, conservation, and in particular modelling, that might impact on the template structure, content and procedural implementation of preregistration.

#### Analysis and Design ####

**1\. Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**

@@

A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on 'good modelling practice' and structured decision-making literature in ecological management. Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substeps', in line with the ontology we specified for translating a modelling workflow into a preregistration template [section link], \<insert link to appendices or just present the final?\>. Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and substeps. We revised the coding structure \<inductively (is this the word)\>, according to patterns and themes identified across multiple personal workflows. For example, we added the final phase \<name for last phase\>, which aimed to capture the fact that analysts usually present the results of their modelling to clients, decision-makers or other stakeholders that must interpret the evidence and are responsible for making management decisions. In this way, we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both \<idealised norms and norms of practice, Pu et al., see other philosophy of sci refs\>. This workflow is used to provide structure and content to the preregistration template.

@@

We translated the model-based workflow identified in part I of workshop analysis into a text-based preregistration template. Reflecting on existing templates \<link to OSF repository of templates page\>, we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part I of workshop output analysis. Following the features of a good guideline identified by Moallemi et al. [-@Moallemi2019], we provide examples of choices or illustrations to clarify each step of the preregistration, for parts of the modelling process that are semantically, epistemically and/or philosophically contentious or ambiguous \<link to paper describing contentious issues in modelling history\>. We synthesised the following ontology to build our preregistration template:

@@

2\. Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps', which corresponds to an activity or set of related activities in a scientific workflow. \<again, these are being scrapped and turned into tool-tips.\>

@@

@@

**What about the challenges and solutions part + the group discussion? Make sure to discuss the atlas TI part for this...**

@@

@@

The modelling workflow + adaptive preregistration process described below articulates the inner logic of the preregistration templates that underpins the content, structure of the template, as well as our proposed procedure and interface for implementing the template.

@@

The template is designed to be used within an adaptive preregistration methodology \<link to section\>.

@@

I think the preregistration can easily be usd with a reporting check
